Bayesian Sparsification Methods for Deep Complex-valued Networks

### 2020-06-23 paper/paper.tex
edited appendix-D on cplx stuff

\begin{figure}[!t]
  \centering
  \includegraphics[width=\columnwidth]{grad_log.pdf}
  \caption{$\tfrac{d f(\alpha)}{d \log{\alpha}}$ of the approximation
  \eqref{eq:improper-kl-div-real-approx}, MC estimate of \eqref{eq:improper-kl-div-real},
  and the exact derivative using \eqref{eq:real-kl-div-deriv-log}.}
  \label{fig:molchanov-derivative-replica}
\end{figure}

@@

% essential intro into Wirtinger calculus (CR)
% https://math.stackexchange.com/a/444493
Wirtinger ($\cplx\real$) calculus relies on the natural identification of $\cplx$ with $
  \real \times \real
$, and regards $
  f\colon \cplx \to \cplx
$ as an equivalent in algebraic sense function $F\colon \real^2 \to \cplx$ defined $
  f(z) = f(u + jv) = F(u, v)
$. Within this framework the differential of $f$ at $z = u + jv \in \cplx$ is
$$
df(z)
  = \frac{\partial f}{\partial z} dz
    + \frac{\partial f}{\partial \conj{z}} d\conj{z}
   % = \frac12\bigl(
   %    \frac{\partial}{\partial u}
   %    - j \frac{\partial}{\partial v}
   % \bigr) F (du - j dv)
   % + \frac12\bigl(
   %    \frac{\partial}{\partial u}
   %    + j \frac{\partial}{\partial v}
   % \bigr) F (du - j dv)
   % = \frac12 \bigl(
   %    \frac{\partial F}{\partial u} du - j \frac{\partial F}{\partial v} du
   %    + j \frac{\partial F}{\partial u} dv + \frac{\partial F}{\partial v} dv
   % \bigr)
   % + \frac12 \bigl(
   %    \frac{\partial F}{\partial u} du + j \frac{\partial F}{\partial v} du
   %    - j \frac{\partial F}{\partial u} dv + \frac{\partial F}{\partial v} dv
   % \bigr)
   = \frac{\partial F}{\partial u} du
     + \frac{\partial F}{\partial v} dv
   = dF(u, v)
  \,, $$
with formally defined derivatives $
  \tfrac{\partial}{\partial z}
    = \tfrac12 \bigl(
      \tfrac{\partial}{\partial u}
      - j \tfrac{\partial}{\partial v}
    \bigr)
$ and $
  \tfrac{\partial}{\partial \conj{z}}
    = \tfrac12 \bigl(
      \tfrac{\partial}{\partial u}
      + j \tfrac{\partial}{\partial v}
    \bigr)
$, and differentials $dz = du + j dv$ and $d\conj{z} = du - j dv$. This implies that the
complex argument and its conjugate are treated as independent variables. Cauchy-Riemann
conditions $
  -j \tfrac{\partial F}{\partial v} = \tfrac{\partial F}{\partial u}
$ can be expressed as $
  \tfrac{\partial f}{\partial \conj{z}} = 0
$ in this notation. Thus $\cplx\real$ calculus subsumes the usual $\cplx$-calculus on
holomorphic functions, when $f(z)$, regarded as $f(z, \conj{z})$, is constant with respect
to $\conj{z}$. The usual rules of calculus, like chain and product rules, follow directly
from the definition of the operators, e.g.
$$
  \frac{\partial (f\circ g)}{\partial z}
    = \frac{\partial f(g(z))}{\partial g} \frac{\partial g(z)}{\partial z}
    + \frac{\partial f(g(z))}{\partial \conj{g}} \frac{\partial \conj{g(z)}}{\partial z}
    % = \nabla G \nabla F
  \,. $$
% Show that $dh(z) = df(g(z)) dg(z)$.
% $h(z) = f(g(z)) = F(G_r(u, v), G_i(u, v)) = H(u, v)$
% $$
% dH
%   = \tfrac{\partial F}{\partial x} \tfrac{\partial G_r}{\partial u} du
%   + \tfrac{\partial F}{\partial x} \tfrac{\partial G_r}{\partial v} dv
%   + \tfrac{\partial F}{\partial y} \tfrac{\partial G_i}{\partial u} du
%   + \tfrac{\partial F}{\partial y} \tfrac{\partial G_i}{\partial v} dv
%   = \tfrac{\partial F}{\partial r} d\Re g(z)
%   + \tfrac{\partial F}{\partial i} d\Im g(z)
%   = df(\Re g(z) + j\Im g(z)) d g(z)
%   = df(g(z)) d g(z)
%   \,. $$

In machine learning tasks the target objective is typically empirical risk, which has to
be real-valued to be minimized. Nevertheless, the expression of the $\cplx\real$ gradient
is compatible with what is expected, when $f$ is treated like a $\real^2$ function. For a
real-valued $f\colon \cplx \to \real$ we have $\conj{f} = f$, which implies $
  \tfrac{\partial f}{\partial \conj{z}}
    = \tfrac{\partial \conj{f}}{\partial \conj{z}}
    = \conj{\tfrac{\partial f}{\partial z}}
$, whence
$$
df
  = \tfrac{\partial f}{\partial z} dz
    + \tfrac{\partial f}{\partial \conj{z}} d\conj{z}
  % = \conj{\tfrac{\partial \conj{f}}{\partial \conj{z}}} dz
  % + \conj{\conj{\tfrac{\partial f}{\partial \conj{z}}} dz}
  = 2 \Re \Bigl(
    \conj{\tfrac{\partial f}{\partial \conj{z}}} dz
  \Bigr)
  = 2 \Re \bigl(
    \tfrac{\partial f}{\partial z} dz
  \bigr)
  \,. $$
Thus, the gradient of $f$ at $z$ is given by $
  \nabla_z f(z)
    = \tfrac{\partial F}{\partial u}
      + j \tfrac{\partial F}{\partial v}
$.

Natural identification of $\cplx$ with $\real\times \real$, storing the real and imaginary
parts in interleaved format, emulation of $\cplx$-algebra using $\real$-valued arithmetic,
and Wirtinger calculus make it possible to reuse $\real$ back-propagation and existing
auto-differentiation frameworks \citep{trabelsi_deep_2018}.

@@

% proof that such operators exist and are unique (well this is an obvious statement)
Consider $L \colon \cplx^m \to \cplx^n$ -- linear in $\cplx$. Then $
  L(u + jv) = L u + j L v
$ for any $u, v \in \real^m$, which implies that the effect of $L$ as $\cplx$-linear
operator is determined by its restriction onto $\real^n$. Let $
  F = L\vert_{\real^m}
  \colon \real^m \to \cplx^n
$ and observe that $F_r = \Re \circ F$ and $F_i = \Im \circ F$ are $\real$-linear operators
such that $F = F_r + j F_i$ (pointwise). Indeed,
$$
  F_r(a + \lambda b)
  % = \Re F(a + \lambda b)
  = \Re L(a + \lambda b)
  % = \Re \bigl( L a + \lambda L b \bigr)
  = \Re L a + \lambda \Re L b
  % = \Re F a + \lambda \Re F b
  = F_r a + \lambda F_r b
  \,. $$
Therefore for any $\cplx$-linear operator $L$ there are $\real$-linear operators $U, V$
such that
$$
L z 
  = (U + j V) (\Re z + j \Im z)
  % = (U + j V) \Re z + j (U + j V) \Im z
  % = U \Re z + j V \Re z + j \bigl(U \Im z + j V \Im z \bigr)
  = U \Re z - V \Im z + j \bigl( V \Re z + U \Im z \bigr)
  % = (U + j V) z
  \,. $$
Uniqueness of these operators follows, if this decomposition is applied to any $z$ with
$\Im z = 0$.

>>>>

\begin{figure}[!t]
  \centering
  \includegraphics[width=\columnwidth]{grad_log.pdf}
  \caption{$\tfrac{d K(\alpha)}{d \log{\alpha}}$ of the approximation
  \eqref{eq:improper-kl-div-real-approx}, MC estimate of \eqref{eq:improper-kl-div-real},
  and the exact derivative using \eqref{eq:real-kl-div-deriv-log}.}
  \label{fig:molchanov-derivative-replica}
\end{figure}

@@

% essential intro into Wirtinger calculus (CR)
% https://math.stackexchange.com/a/444493
Wirtinger ($\cplx\real$) calculus relies on the natural identification of $\cplx$ with $
  \real^2
$, and regards $
  f\colon \cplx \to \cplx
$ as an algebraically equivalent function $F\colon \real^2 \to \cplx$ defined $
  f(z) = f(u + \iu v) = F(u, v)
$. It enables general treatment of functions of vector $\cplx$-argument that possess partial
derivatives with respect to real and imaginary parts, yet do are not required to satisfy
Cauchy-Riemann conditions. In $\cplx\real$ calculus the complex argument $z$ and its
conjugate $\conj{z}$ act as independent variables and $f(z)$ is treated as $f(z, \conj{z})$
by way of geometric transformations $z = u + \iu v$ and $\conj{z} = u - \iu v$.

Wirtinger partial derivative operators are formally defined as $
  \tfrac{\partial}{\partial z}
    = \tfrac12 \bigl(
      \tfrac{\partial}{\partial u}
      - \iu \tfrac{\partial}{\partial v}
    \bigr)
  \tfrac{\partial}{\partial \conj{z}}
    = \tfrac12 \bigl(
      \tfrac{\partial}{\partial u}
      + \iu \tfrac{\partial}{\partial v}
    \bigr)
$ and differentials are $dz = du + \iu dv$ and $d\conj{z} = du - \iu dv$. In this paradigm
The usual rules of calculus, like chain and product rules, follow directly
from the definition of the operators, e.g.
\begin{equation}
    \frac{\partial (f\circ g)}{\partial z}
    = \frac{\partial f(g(z))}{\partial g} \frac{\partial g(z)}{\partial z}
    + \frac{\partial f(g(z))}{\partial \conj{g}} \frac{\partial \conj{g(z)}}{\partial z}
    % = \nabla G \nabla F
  \,.  \notag
The total differential of $f$ at $z = u + \iu v \in \cplx$ is
\begin{eqnarray}  
df(z)
  &=& \frac{\partial f}{\partial z} dz
    + \frac{\partial f}{\partial \conj{z}} d\conj{z}
  % &=& \frac12\biggl(
  %    \frac{\partial F}{\partial u}
  %    - \iu \frac{\partial F}{\partial v}
  % \biggr) (du + \iu dv)
  % + \frac12\biggl(
  %    \frac{\partial F}{\partial u}
  %    + \iu \frac{\partial F}{\partial v}
  % \biggr) (du - \iu dv)
  &=& \frac12 \biggl(
     \frac{\partial F}{\partial u} du
     - \cancel{\iu \frac{\partial F}{\partial v} du}
     + \cancel{\iu \frac{\partial F}{\partial u} dv}
     + \frac{\partial F}{\partial v} dv
  + \frac12 \biggl(
     \frac{\partial F}{\partial u} du
     + \cancel{\iu \frac{\partial F}{\partial v} du}
     - \cancel{\iu \frac{\partial F}{\partial u} dv}
     + \frac{\partial F}{\partial v} dv
  % &=& \frac{\partial F}{\partial u} du
  %    + \frac{\partial F}{\partial v} dv
  &=& dF(u, v)
  \,,  \notag
% \begin{eqnarray}
%   d(f\circ g)
%     &=&
%       \frac{\partial f(g(z))}{\partial g}
%       dg
%     + \frac{\partial f(g(z))}{\partial \conj{g}}
%       d\conj{g}
%     &=&
%       \frac{\partial f(g(z))}{\partial g}
%       \biggl\{
%         \frac{\partial g}{\partial z}
%         dz
%       + \frac{\partial g}{\partial \conj{z}}
%         d\conj{z}
%       \biggr\}
%     + \frac{\partial f(g(z))}{\partial \conj{g}}
%       \biggl\{
%         \frac{\partial \conj{g}}{\partial z}
%         dz
%       + \frac{\partial \conj{g}}{\partial \conj{z}}
%         d\conj{z}
%       \biggr\}
%     &=&
%         \frac{\partial f(g(z))}{\partial g} \frac{\partial g}{\partial z}
%         dz
%       + \frac{\partial f(g(z))}{\partial g} \frac{\partial g}{\partial \conj{z}}
%         d\conj{z}
%       + \frac{\partial f(g(z))}{\partial \conj{g}} \frac{\partial \conj{g}}{\partial z}
%         dz
%       + \frac{\partial f(g(z))}{\partial \conj{g}} \frac{\partial \conj{g}}{\partial \conj{z}}
%         d\conj{z}
%     &=&
%       \biggl\{
%         \frac{\partial f(g(z))}{\partial g} \frac{\partial g}{\partial z}
%         + \frac{\partial f(g(z))}{\partial \conj{g}} \conj{\frac{\partial g}{\partial \conj{z}}}
%       \biggr\} dz
%       + \biggl\{
%         \frac{\partial f(g(z))}{\partial g} \frac{\partial g}{\partial \conj{z}}
%         + \frac{\partial f(g(z))}{\partial \conj{g}} \conj{\frac{\partial g}{\partial z}}
%       \biggr\}
%         d\conj{z}
%     \,.
% \end{eqnarray}
% Show that $dh(z) = df(g(z)) dg(z)$.
% $h(z) = f(g(z)) = F(G_r(u, v), G_i(u, v)) = H(u, v)$
% dH
%   = \tfrac{\partial F}{\partial x} \tfrac{\partial G_r}{\partial u} du
%   + \tfrac{\partial F}{\partial x} \tfrac{\partial G_r}{\partial v} dv
%   + \tfrac{\partial F}{\partial y} \tfrac{\partial G_i}{\partial u} du
%   + \tfrac{\partial F}{\partial y} \tfrac{\partial G_i}{\partial v} dv
%   = \tfrac{\partial F}{\partial r} d\Re g(z)
%   + \tfrac{\partial F}{\partial i} d\Im g(z)
%   = df(\Re g(z) + j\Im g(z)) d g(z)
%   = df(g(z)) d g(z)
At the same time the Cauchy-Riemann conditions $
  -\iu \tfrac{\partial F}{\partial v} = \tfrac{\partial F}{\partial u}
$ can be expressed as $
  \tfrac{\partial f}{\partial \conj{z}} = 0
$. Thus $\cplx\real$ calculus subsumes the usual $\cplx$-calculus of holomorphic functions,
since $f(z) = f(z, \conj{z})$ is constant with respect to $\conj{z}$ in the latter.

In optimization-related tasks the objective is $
  f\colon \cplx\to \real
$, meaning that if it were to satisfy the Cauchy-Riemann conditions, then it necessarily
should have been constant. Nevertheless, the expression of the $\cplx\real$ gradient is
compatible with what is expected, when $f$ is treated like a $\real^2$ function. For such
$f$ we have $\conj{f} = f$, which implies $
  \tfrac{\partial f}{\partial \conj{z}}
    = \tfrac{\partial \conj{f}}{\partial \conj{z}}
    = \conj{\tfrac{\partial f}{\partial z}}
$, whence
\begin{equation*}
  df
    = \tfrac{\partial f}{\partial z} dz
      + \tfrac{\partial f}{\partial \conj{z}} d\conj{z}
    % = \tfrac{\partial f}{\partial z} dz
    %   + \conj{\tfrac{\partial f}{\partial z}} d\conj{z}
    = \tfrac{\partial f}{\partial z} dz
      + \conj{\tfrac{\partial f}{\partial z} dz}
    % = 2 \Re
    %   \Bigl(
    %     \conj{\tfrac{\partial f}{\partial \conj{z}}} dz
    %   \Bigr)
    = 2 \Re
      \bigl(
        \tfrac{\partial f}{\partial z} dz
      \bigr)
    \,.
\end{equation*}
Therefore the gradient the direction of greatest ascent of $f$ at $z$ is given by $
  \nabla_{\conj{z}} f(z)
    = \conj{\tfrac{\partial f}{\partial z}}
    = \tfrac{\partial F}{\partial u}
    + \iu \tfrac{\partial F}{\partial v}
$.
Natural identification $\cplx \simeq \real^2$, backed by Wirtinger calculus, and emulation
of $\cplx$-arithmetic in computational graphs with $\real$-valued operations makes it
possible to reuse $\real$ back-propagation and existing auto-differentiation frameworks.

@@

% proof that such operators exist and are unique (well this is an obvious statement)
Consider $L \colon \cplx^m \to \cplx^n$ -- linear in $\cplx$. Then $
  L(u + jv) = L u + \iu L v
$ for any $u, v \in \real^m$, which implies that the effect of $L$ as $\cplx$-linear
operator is determined by its restriction onto $\real^n$. Let $
  F = L\vert_{\real^m}
  \colon \real^m \to \cplx^n
$ and observe that $F_r = \Re \circ F$ and $F_i = \Im \circ F$ are $\real$-linear operators
such that $F = F_r + \iu F_i$ (pointwise). Indeed,
$$
  F_r(a + \lambda b)
  % = \Re F(a + \lambda b)
  = \Re L(a + \lambda b)
  % = \Re \bigl( L a + \lambda L b \bigr)
  = \Re L a + \lambda \Re L b
  % = \Re F a + \lambda \Re F b
  = F_r a + \lambda F_r b
  \,. $$
Therefore for any $\cplx$-linear operator $L$ there are $\real$-linear operators $U, V$
such that
$$
L z 
  = (U + \iu V) (\Re z + \iu \Im z)
  % = (U + \iu V) \Re z + \iu (U + \iu V) \Im z
  % = U \Re z + \iu V \Re z + \iu \bigl(U \Im z + \iu V \Im z \bigr)
  = U \Re z - V \Im z + \iu \bigl( V \Re z + U \Im z \bigr)
  % = (U + \iu V) z
  \,. $$
Uniqueness of these operators follows, if this decomposition is applied to any $z$ with
$\Im z = 0$.
